---
title: "VIIRS analysis"
output: html_document
---

# Setup

```{r}
library(ggplot2)
library(dplyr)

# Establish connection to BigQuery project
con <- DBI::dbConnect(
  drv = bigrquery::bigquery(),
  project = "world-fishing-827",
  use_legacy_sql = FALSE
)
options(scipen = 20)

bigrquery::bq_auth(email = "masaki@globalfishingwatch.org")



round1000 <- function(x){
  round(x*(10^-6),3)*10^6
}


round0.1 <- function(x){
  round(x*10)/10
}
```

# Create tables

```{r, eval=FALSE}

# Create a table that contains cloud mask within squid fishing areas 
query1 <- readr::read_file(here::here("queries", "VIIRS", "01_create_cloudmask_squid_area.sql"))

# Create a table that contains VIIRS-AIS matching data within squid fishing areas 
query2 <- readr::read_file(here::here("queries", "VIIRS", "02_create_viirs_matching_squid_area_no_overlap_local_night_2017_2021.sql"))

# Create a table that contains all the AIS messages within squid fishing areas
query3 <- readr::read_file(here::here("queries", "VIIRS", "03_create_pipe_in_the_squid_area.sql"))

# Create a table that contains daily night loitering information 
# for squid vessels in the squid fishing areas
query4 <- readr::read_file(here::here("queries", "VIIRS", "04_create_nightloitering_vessel_points_only_squid_local_night.sql"))
```






# GET VIIRS and AIS count

```{r}

# Extracting only VBDs with radiance greater than RAD_THRESHOLD
RAD_THRESHOLD <- 10

# Extracting only night loitering vessels with daily night loitering hours greater than NL_HOUR_THRESHOLD

NL_HOUR_THRESHOLD <- 1

# VIIRS-AIS matching with score greater than SCORE_THRESHOLD regarded as reliable matching
SCORE_THRESHOLD <- 0.01


# Get daily counts of VIIRS detentions and night loitering vessels
query <- glue::glue(readr::read_file(  here::here("queries", "VIIRS", "05_get_viirs_and_ais_counts_in_squid_area_only_squid_ssvid.sql")))



viirs_ais_counts_orig_df <-
  DBI::dbGetQuery(con, query) %>%
  mutate(
    score_threshold = SCORE_THRESHOLD,
    nl_threshold = NL_HOUR_THRESHOLD,
    rad_threshold = RAD_THRESHOLD
  )


rds_file <- "viirs_ais_counts_for_paper.rds"
saveRDS(viirs_ais_counts_orig_df, rds_file)
```


# Read AIS and VIIRS


```{r}
rds_file <- "viirs_ais_counts_for_paper.rds"
viirs_ais_counts_orig_df <- readRDS(rds_file)
```


# eliminate anomaly date from counting data


```{r}
# eliminate anomaly date
viirs_ais_counts_df <-
  viirs_ais_counts_orig_df %>%
  rename(year_month0.5 = year_half_month) %>%
  filter(!(area == "se_pacific" & date == as.Date("2017-04-07")))


# check data after eliminating the anomaly date
# ggplot(viirs_ais_counts_df) +
#   geom_bar(aes(date, count_viirs_total),
#            stat = "identity",
#            width = 1) +
#   #geom_line(aes(date, count_viirs_total_halfmonth), width = 1)+
#   facet_wrap( ~ area) +
#   ggsave(
#     "viirs_ais_counts_df.png",
#     width = 20,
#     height = 20,
#     units = "cm"
#   )
```


# Extract the date when the maximum number of VIIRS detections was observed every 15 days.


```{r}

#########################
# Extract AIS/VIIRS on the date where maximum VIIRS are observed for each 15 days
suppressWarnings(
  viirs_ais_on_15days_max_df <-
    viirs_ais_counts_df %>%
    group_by(year_month0.5, area) %>%
    mutate(max_count_half_month = max(count_viirs_total, na.rm = TRUE)) %>%
    filter(count_viirs_total == max_count_half_month) %>%
    arrange(area, date) %>%
    mutate(rownum = row_number()) %>%
    filter(rownum == 1)
)

# Change column name and select only necessary columns
viirs_ais_half_month_df <-
  viirs_ais_on_15days_max_df %>%
  rename(
    count_viirs_total_half_month = count_viirs_total,
    count_viirs_matched_half_month = count_viirs_matched,
    # count night loitering vessels
    count_nl_half_month = count_nl,
    count_nl_matched_half_month = count_nl_matched,
  ) %>%
  select(
    area,
    year_month0.5,
    count_viirs_total_half_month,
    count_viirs_matched_half_month,
    count_nl_half_month,
    count_nl_matched_half_month,
  ) %>%
  arrange(area, year_month0.5)

# Join 15-day max count to every day count
viirs_ais_counts_daily_month0.5_df <-
  viirs_ais_counts_df %>%
  left_join(viirs_ais_half_month_df, by = c("area", "year_month0.5"))
```



# Calculate the detection rate of AIS vessels



```{r}
# ######################
# detection_rate_by_month3_df <-
#   viirs_ais_on_15days_max_df %>%
#   group_by(year_month3, area) %>%
#   summarise(
#     count_viirs_total_month3 = sum(count_viirs_total, na.rm = TRUE),
#     count_viirs_matched_month3 = sum(count_viirs_matched, na.rm = TRUE),
#     count_nl_month3 = sum(count_nl, na.rm = TRUE),
#     count_nl_matched_month3 = sum(count_nl_matched, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   mutate(detection_rate_by_month3 = count_nl_matched_month3 / count_nl_month3) %>%
#   select(year_month3,
#          area,
#          detection_rate_by_month3,) %>%
#   arrange(area, year_month3)
# 
# # ######################
# detection_rate_by_month4_df <-
#   viirs_ais_on_15days_max_df %>%
#   group_by(year_month4, area) %>%
#   summarise(
#     count_viirs_total_month4 = sum(count_viirs_total, na.rm = TRUE),
#     count_viirs_matched_month4 = sum(count_viirs_matched, na.rm = TRUE),
#     count_nl_month4 = sum(count_nl, na.rm = TRUE),
#     count_nl_matched_month4 = sum(count_nl_matched, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   mutate(detection_rate_by_month4 = count_nl_matched_month4 / count_nl_month4) %>%
#   select(year_month4,
#          area,
#          detection_rate_by_month4) %>%
#   arrange(area, year_month4)
# 
# 
# 
# #######################
# detection_rate_by_month6_df <-
#   viirs_ais_on_15days_max_df %>%
#   group_by(year_month6, area) %>%
#   summarise(
#     count_viirs_total_month6 = sum(count_viirs_total, na.rm = TRUE),
#     count_viirs_matched_month6 = sum(count_viirs_matched, na.rm = TRUE),
#     count_nl_month6 = sum(count_nl, na.rm = TRUE),
#     count_nl_matched_month6 = sum(count_nl_matched, na.rm = TRUE),
#     .groups = "drop"
#   ) %>%
#   mutate(detection_rate_by_month6 = count_nl_matched_month6 / count_nl_month6) %>%
#   select(year_month6,
#          area,
#          detection_rate_by_month6) %>%
#   arrange(area, year_month6)



# Join daily AIS/VIIRS and 
# viirs_ais_counts_daily_month0.5_with_detection_rates_df <-
#   viirs_ais_counts_daily_month0.5_df %>%
#   left_join(detection_rate_by_month3_df,
#             by = c("year_month3", "area")) %>%
#   left_join(detection_rate_by_month4_df,
#             by = c("year_month4", "area")) %>%
#   left_join(detection_rate_by_month6_df,
#             by = c("year_month6", "area"))


viirs_ais_counts_daily_month0.5_with_detection_rates_df <-
   viirs_ais_counts_daily_month0.5_df
```


# Estimate dark vessels

Estimate total number of vessels from the detection rate of AIS vessels


```{r}
#########################
# calculate dark vessels
# viirs_ais_dark_df <-
#   viirs_ais_counts_daily_month0.5_with_detection_rates_df %>%
#   # estimate the total number of vessels by using 6month detection rate of night loiterring vessels
#   mutate(
#     estimated_total_vessels_month3 = count_viirs_total_half_month / detection_rate_by_month3,
#     estimated_total_vessels_month4 = count_viirs_total_half_month / detection_rate_by_month4,
#     estimated_total_vessels_month6 = count_viirs_total_half_month / detection_rate_by_month6
#   )  %>%
#   # add average number of night loitering AIS vessels
#   # for each 15 days
#   group_by(area, year_month0.5) %>%
#   mutate(count_nl_half_month_mean = mean(count_nl, na.rm = TRUE)) %>%
#   ungroup()



viirs_ais_dark_df <-
  viirs_ais_counts_daily_month0.5_with_detection_rates_df
```




# Plot for checking

## daily plot

```{r}
daily_area_long_df <-
  viirs_ais_dark_df %>%
  mutate(
    #estimated_total_vessels = estimated_total_vessels_month6,
    total_viirs = count_viirs_total,
    total_ais = count_nl,
    viirs_ais_matched = count_nl_matched,
    only_viirs = total_viirs - viirs_ais_matched,
    #only_ais = total_ais - viirs_ais_matched,
    #total_ais_viirs = viirs_ais_matched + only_viirs + only_ais,
  ) %>%
  mutate(
    area = case_when(
      area == "nw_indian" ~ "NW Indian",
      area == "nw_pacific" ~ "NW Pacific",
      area == "se_pacific" ~ "SE Pacific",
      area == "sw_atlantic" ~ "SW Atlantic"
    )
  ) %>%
  mutate(year = lubridate::year(date)) %>%
  select(
    date,
    area,
    only_viirs,
    #only_ais,
    viirs_ais_matched,
    #only_estimate,
  ) %>% 
  tidyr::pivot_longer(
    cols = c(
      only_viirs,
      #only_ais,
      viirs_ais_matched,
      #only_estimate
    ),
    names_to = "Source",
    values_to = "Vessel count",
    
  ) %>% 
  mutate(Source = factor(Source, levels = rev(
    c(
      "viirs_ais_matched",
      #"only_ais",
      "only_viirs"
      #"only_estimate"
    )
  )))%>%
  mutate(`Vessel count` = ifelse(is.na(`Vessel count`), 0, `Vessel count`)) %>% 
  mutate(Source = case_when(
      Source == "viirs_ais_matched" ~ "Detected by both VIIRS and AIS",
      #Source == "only_ais" ~ "Detected only by AIS",
      Source == "only_viirs" ~ "Detected only by VIIRS"
      #Source == "only_estimate" ~ "Potentially undetected vessels"
    
  )) %>% 
  mutate(Source = factor(
    Source,
    levels = c(
      "Detected only by VIIRS",
      "Detected by both VIIRS and AIS"
      #"Potentially undetected vessels",
      #"Detected only by AIS",
      
    )
  )) %>%
  mutate(`Detection type` = Source)



plot <- 

ggplot(daily_area_long_df) +
  geom_bar(aes(date, `Vessel count`, fill = Source),
           stat = "identity",
           width = 1) +
  fishwatchr::theme_gfw() +
  facet_wrap(area ~ .) +
  xlim(as.Date("2017-01-01"), as.Date("2021-12-31")) +
  scale_fill_manual(
    values = c(
      #"only_ais" = "blue",
      "Detected only by VIIRS" = fishwatchr::gfw_palette("chart")[6],
      "Detected by both VIIRS and AIS" = fishwatchr::gfw_palette("chart")[1]
      #"only_estimate" = "gray"
    )
  ) +
  scale_x_date(
    breaks = seq.Date(as.Date("2017-01-01"), as.Date("2021-01-01"), by = "year"),
    labels = seq(2017, 2021, by = 1)
  ) 


  ggsave(filename="vessel_count_daily_by_region.png", plot=plot, width = 20, height = 15, units = "cm", dpi = 300)
```



## half-monthly count

This plot is only for checking

```{r}
half_monthly_area_long_df <-
  viirs_ais_dark_df %>%
  mutate(
    #estimated_total_vessels = estimated_total_vessels_month6,
    total_viirs = count_viirs_total_half_month,
    total_ais = count_nl_half_month,
    viirs_ais_matched = count_nl_matched_half_month,
    only_viirs = total_viirs - viirs_ais_matched,
    #only_ais = total_ais - viirs_ais_matched,
    #total_ais_viirs = viirs_ais_matched + only_viirs + only_ais,
  ) %>%
  mutate(
    area = case_when(
      area == "nw_indian" ~ "NW Indian",
      area == "nw_pacific" ~ "NW Pacific",
      area == "se_pacific" ~ "SE Pacific",
      area == "sw_atlantic" ~ "SW Atlantic"
    )
  ) %>%
  mutate(year = lubridate::year(date)) %>%
  select(
    date,
    area,
    only_viirs,
    #only_ais,
    viirs_ais_matched,
    #only_estimate,
  ) %>% 
  tidyr::pivot_longer(
    cols = c(
      only_viirs,
      #only_ais,
      viirs_ais_matched,
      #only_estimate
    ),
    names_to = "Source",
    values_to = "Vessel count",
    
  ) %>% 
  mutate(Source = factor(Source, levels = rev(
    c(
      "viirs_ais_matched",
      #"only_ais",
      "only_viirs"
      #"only_estimate"
    )
  )))%>%
  mutate(`Vessel count` = ifelse(is.na(`Vessel count`), 0, `Vessel count`)) %>% 
  mutate(Source = case_when(
      Source == "viirs_ais_matched" ~ "Detected by both VIIRS and AIS",
      #Source == "only_ais" ~ "Detected only by AIS",
      Source == "only_viirs" ~ "Detected only by VIIRS"
      #Source == "only_estimate" ~ "Potentially undetected vessels"
    
  )) %>% 
  mutate(Source = factor(
    Source,
    levels = c(
      "Detected only by VIIRS",
      "Detected by both VIIRS and AIS"
      #"Potentially undetected vessels",
      #"Detected only by AIS",
      
    )
  )) %>%
  mutate(`Detection type` = Source)


plot <- 
ggplot(half_monthly_area_long_df) +
  geom_bar(aes(date, `Vessel count`, fill = Source),
           stat = "identity",
           width = 1) +
  fishwatchr::theme_gfw() +
  facet_wrap(area ~ .) +
  xlim(as.Date("2017-01-01"), as.Date("2021-12-31")) +
  scale_fill_manual(
    values = c(
      #"only_ais" = "blue",
      "Detected only by VIIRS" = fishwatchr::gfw_palette("chart")[6],
      "Detected by both VIIRS and AIS" = fishwatchr::gfw_palette("chart")[1]
      #"only_estimate" = "gray"
    )
  ) +
  scale_x_date(
    breaks = seq.Date(as.Date("2017-01-01"), as.Date("2021-01-01"), by = "year"),
    labels = seq(2017, 2021, by = 1)
  )






ggsave(
  filename = "vessel_count_half_month_by_region.png",
  plot = plot,
  width = 20,
  height = 15,
  units = "cm",
  dpi = 300
)

```



# Figure 2

This plot is used for the Figure 2

```{r}

yearly_area_long_df <-
  viirs_ais_dark_df %>%
  mutate(
    #estimated_total_vessels = estimated_total_vessels_month6,
    total_viirs = count_viirs_total_half_month,
    #total_ais = count_nl_half_month,
    viirs_ais_matched = count_nl_matched_half_month,
    only_viirs = total_viirs - viirs_ais_matched,
    #only_ais = total_ais - viirs_ais_matched,
    #total_ais_viirs = viirs_ais_matched + only_viirs + only_ais,
  ) %>%
  mutate(
    area = case_when(
      area == "nw_indian" ~ "NW Indian",
      area == "nw_pacific" ~ "NW Pacific",
      area == "se_pacific" ~ "SE Pacific",
      area == "sw_atlantic" ~ "SW Atlantic"
    )
  ) %>%
  # mutate(only_estimate =
  #          ifelse(
  #            estimated_total_vessels > total_ais_viirs,
  #            round(estimated_total_vessels - total_ais_viirs),
  #            0
  #          )) %>%
  select(date,
         area,
         only_viirs,
         #only_ais,
         viirs_ais_matched,
         #only_estimate,
         ) %>%
  tidyr::pivot_longer(
    cols = c(only_viirs,
             #only_ais,
             viirs_ais_matched,
             #only_estimate
             ),
    names_to = "Source",
    values_to = "Vessel count",
    ) %>% 
   mutate(Source = factor(Source, levels = rev(c(
     "viirs_ais_matched",
     #"only_ais",
     "only_viirs"
     #"only_estimate"
   )))) %>%
  mutate(`Vessel count` = ifelse(is.na(`Vessel count`), 0, `Vessel count`)) %>%
  mutate(
    Source = case_when(
      Source == "viirs_ais_matched" ~ "Detected by both VIIRS and AIS",
      #Source == "only_ais" ~ "Detected only by AIS",
      Source == "only_viirs" ~ "Detected only by VIIRS"
      #Source == "only_estimate" ~ "Potentially undetected vessels"
    )
  ) %>% 
  mutate(Source = factor(
    Source,
    levels = c("Detected by both VIIRS and AIS",
               "Detected only by VIIRS"
               #"Potentially undetected vessels",
               #"Detected only by AIS"
               )
    )) %>% 
  mutate(`Detection type` = Source) %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(area, year, Source) %>%
  summarise(`Vessel count` = sum(`Vessel count`, na.rm = TRUE)) %>%
  group_by(area, year) %>%
  mutate(total = sum(`Vessel count`, na.rm = TRUE),
         percentage = round(100 * (`Vessel count` / total))) %>%
  mutate(percentage_display = ifelse(
    Source == "Detected by both VIIRS and AIS",
    "",
    paste0(percentage, "%")
  ))
               
               
plot <-           
ggplot(yearly_area_long_df) +
  geom_bar(aes(year, `Vessel count`, fill = Source),
           stat = "identity",
           width = 0.9) +
  geom_text(aes(x = year, y = `Vessel count` * 0.5, label = percentage_display),
            size = 2) +
  fishwatchr::theme_gfw() +
  facet_wrap(area ~ ., strip.position = "top") +
  #xlim(as.Date("2017-01-01"), as.Date("2021-12-31")) +
  scale_fill_manual(
    values = c(
      #"only_ais" = "blue",
      "Detected only by VIIRS" = fishwatchr::gfw_palette("chart")[6],
      "Detected by both VIIRS and AIS" = fishwatchr::gfw_palette("chart")[1]
      #"only_estimate" = "gray"
    )
  ) +
  #scale_x_date(breaks = seq.Date(as.Date("2017-01-01"), as.Date("2021-01-01"), by = "year"), labels = seq(2017,2021,by=1))+
  xlab("Year") +
  ylab("Vessel days") +
  theme(legend.position="bottom") 



  ggsave(
    here::here("outputs", "figures", "VIIRS", "Figure2.png"),
    plot=plot,
    width = 18,
    height = 13,
    units = "cm",
    dpi = 300
  )

```


# Table 1

This CSV data is used for the Table 1

```{r}

# aggregation by region
regional_df <-
  yearly_area_long_df %>%
  tidyr::pivot_wider(
    id_cols = c(year, area),
    names_from = Source,
    values_from = `Vessel count`
  ) %>%
  mutate(`Vessel Count` = `Detected only by VIIRS` + `Detected by both VIIRS and AIS`) %>%
  mutate(
    `%UNMATCHED` = round0.1(100 * `Detected only by VIIRS` / `Vessel Count`),
    `%MATCHED` = round0.1(100 * `Detected by both VIIRS and AIS` / `Vessel Count`)
  )


# global aggregation
global_df <-
  regional_df %>%
  group_by(year) %>%
  summarise(
    `Vessel Count` = sum(`Vessel Count`),
    `Detected only by VIIRS` = sum(`Detected only by VIIRS`),
    `Detected by both VIIRS and AIS` = sum(`Detected by both VIIRS and AIS`)
  ) %>%
  mutate(
    area  = "Total",
    `%UNMATCHED` = round0.1(100 * `Detected only by VIIRS` / `Vessel Count`),
    `%MATCHED` = round0.1(100 * `Detected by both VIIRS and AIS` / `Vessel Count`)
  )


# conbine regional and global
all_df <-
  bind_rows(regional_df, global_df) %>%
  rename(Region=area)



# Total by region 
total_by_region_df <-
  regional_df %>%
  group_by(area) %>%
  summarise(
    `Vessel Count` = sum(`Vessel Count`),
    `Detected only by VIIRS` = sum(`Detected only by VIIRS`),
    `Detected by both VIIRS and AIS` = sum(`Detected by both VIIRS and AIS`)
  ) %>%
  mutate(
    `%UNMATCHED` = round0.1(100 * `Detected only by VIIRS` / `Vessel Count`),
    `%MATCHED` = round0.1(100 * `Detected by both VIIRS and AIS` / `Vessel Count`)
  ) 



# Total all region all year
total_df <-
  total_by_region_df %>%
  summarise(
    `Vessel Count` = sum(`Vessel Count`),
    `Detected only by VIIRS` = sum(`Detected only by VIIRS`),
    `Detected by both VIIRS and AIS` = sum(`Detected by both VIIRS and AIS`)
  ) %>%
  mutate(
    `%UNMATCHED` = round0.1(100 * `Detected only by VIIRS` / `Vessel Count`),
    `%MATCHED` = round0.1(100 * `Detected by both VIIRS and AIS` / `Vessel Count`)
  ) 



# total for each region and year
display01_df <-
  all_df %>%
  mutate(`Vessel Count` = round1000(`Vessel Count`)) %>% 
  mutate(text = paste0(formatC(`Vessel Count`, format = "d", big.mark = ",")," (",`%UNMATCHED`,"%)")) %>%
  tidyr::pivot_wider(id_cols = Region,
                     names_from = year,
                     values_from = text) 


# total by region over 2017-2020
display02_df <- 
  total_by_region_df %>% 
  rename(Region=area) %>% 
  mutate(`Vessel Count` = round1000(`Vessel Count`)) %>% 
    mutate(Total = paste0(formatC(`Vessel Count`, format = "d", big.mark = ",")," (",`%UNMATCHED`,"%)")) %>% 
  select(Region, Total)

# total over all region all year
display03_df <- 
  total_df %>% 
  mutate(`Vessel Count` = round1000(`Vessel Count`)) %>%
  mutate(Region = "Total",
         Total = paste0(formatC(`Vessel Count`, format = "d", big.mark = ",")," (",`%UNMATCHED`,"%)")) %>% 
  select(Region, Total)


# conbine all the totals
display_df <-
  left_join(display01_df,
            bind_rows(display02_df, display03_df),
            by = "Region")




readr::write_csv(display_df, here::here("outputs", "tables", "VIIRS", "Table_1.csv"))
```


# The value for text

The value used in the text.

The rate of increase in the number of VIIRS detection from 2017 to 2020


```{r}
all_df %>% 
  tidyr::pivot_wider(id_cols = Region , names_from = year, values_from = `Vessel Count`) %>% 
  mutate(ratio_2020_2017 =  `2020` / `2017`)
```

# maximum daily vessel count in the NW Indian


```{r}
half_monthly_area_long_df %>% 
  filter(area=="NW Indian") %>% 
  mutate(year=lubridate::year(date)) %>% 
  group_by(year) %>% 
  summarise(max_count = max(`Vessel count`))
```







