---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(sf)
library(dplyr)
library(ggplot2)
library(fishwatchr)
figure_version = 'v20230113'
```


```{r}
squid_regions_and_eezs <- sf::read_sf(here::here('data','geodata','simple_merged_eez_regions_final_v20220525.geojson'))
                                      
                                      
```

```{r}
(final_reg_regions_map <- ggplot() +
  geom_gfw_outline(center = 160) +
  geom_sf(data = squid_regions_and_eezs %>%
            mutate(label = ifelse(label == 'NATIONAL','EEZ', label)) %>%
            fishwatchr::recenter_sf(center = 160), 
          aes(fill = label, 
              color = label),
          size = 1.5) +
  geom_gfw_land(center = 160) +
  scale_fill_manual('', values = c('#204280','#F8BA47','#EE6256')) +
  scale_color_manual('', values = c('#204280','#F8BA47','#EE6256')) +
  theme_gfw_map(theme = 'light'))

  save_plots(final_reg_regions_map,
           file_name = glue::glue('final_reg_regions_map_{figure_version}'),
           plot_width = 200,
           plot_height = 120,
           units = 'mm',
           dpi = 600,
           figure_version = figure_version,
  output_type = 'png')
```



# Make sure to pull this out later
Put in the data loading section and potentially a config.yaml
```{r}
squid_regulatory_positions <- fishwatchr::gfw_query(
  'SELECT * FROM paper_global_squid.squid_regulatory_positions_eez_v20220525',
  save_query = FALSE,
  run_query = TRUE,
  con = con
)$data

regulatory_hs_list <- squid_regulatory_positions %>%
  #filter(region != 'NATIONAL') %>%
  distinct(ssvid)

connections_hs_list <- squid_fleet_connections_2 %>%
  distinct(ssvid) 

squid_regulatory_positions %>%
  filter(ssvid %in% c(regulatory_hs_list %>%
           filter(!ssvid %in% connections_hs_list$ssvid) %>%
           pull(ssvid))) %>%
  count(ssvid, region)

squid_regulatory_positions %>%
  mutate(flag = ifelse(is.na(flag), 'UNK', flag)) %>%
  #filter(region != 'NATIONAL') %>%
  distinct(ssvid, flag) %>%
  count(flag, sort = TRUE)
  
# squid_regulatory_positions %>%
#   filter(!flag %in% c('CHN','TWN','JPN','KOR')) %>%
#   group_by(ssvid, flag, region) %>%
#   summarize(total_hours = sum(regional_total_hours, na.rm = TRUE),
#             total_positions = sum(regional_positions, na.rm = TRUE),.groups = 'drop') %>%
#   mutate(region = ifelse(region == 'NATIONAL', 'NATIONAL', 'HIGH SEAS')) %>%
#   group_by(flag, region) %>%
#   summarize(total_hours = sum(total_hours, na.rm = TRUE),
#             total_positions = sum(total_positions, na.rm = TRUE),.groups = 'drop') %>%
#   group_by(flag) %>%
#   mutate(total_hours_all = sum(total_hours),
#          fraction_high_seas = )

squid_regulatory_positions %>%
  filter(flag %in% c('CHN','TWN','JPN','KOR')) %>%
  group_by(ssvid, flag, region) %>%
  summarize(total_hours = sum(regional_total_hours, na.rm = TRUE),
            total_positions = sum(regional_positions, na.rm = TRUE),.groups = 'drop') %>%
  mutate(region = ifelse(region == 'NATIONAL', 'NATIONAL', 'HIGH SEAS')) %>%
  group_by(flag, region) %>%
  summarize(total_hours = sum(total_hours, na.rm = TRUE),
            total_positions = sum(total_positions, na.rm = TRUE),.groups = 'drop')
```

```{r}
squid_arg_diff_positions <- fishwatchr::gfw_query(
  'SELECT * FROM paper_global_squid.squid_argentina_diff_positions_v20220525',
  save_query = FALSE,
  run_query = TRUE,
  con = con
)$data

```

### Using Argentina Boundary

Get the number of hours that were mistakenly (according to Argentina) included
within the EEZ (NATIONAL) that are actually in the high seas (UNREGULATED)

```{r}
hours_in_arg_gap <- squid_arg_diff_positions %>%
  filter(flag %in% c('TWN','JPN','KOR','CHN')) %>%
  summarize(total_hours = sum(regional_total_hours, na.rm = TRUE),
            total_fishing_hours = sum(regional_fishing_hours, na.rm = TRUE),
            .groups = 'drop')
hours_in_arg_gap
# total_hours = 91177.01
# total_fishing_hours = 33928.9
```

## Calculate hours of effort in regions 
Adjusted is adjusted for Argentina EEZ (less national fishing, more unregulated)
```{r}
squid_regulatory_positions_filtered <- squid_regulatory_positions %>%
  filter(flag %in% c('JPN','KOR','TWN','CHN')) %>%
  filter(region %in% c('NATIONAL','RFMO MANAGED','UNREGULATED')) #%>%
  # mutate(region = ifelse(!is.na(is_npfc_regulated) & region == 'NATIONAL', 'RFMO REGULATED', region))

squid_regulatory_positions_filtered %>%
  group_by(region) %>%
  summarize(total_hours = sum(regional_total_hours),
            total_fishing_hours = sum(regional_fishing_hours)) %>%
    ungroup() %>%
    mutate(total_hours_adj = case_when(region == 'NATIONAL' ~ total_hours - hours_in_arg_gap$total_hours,
                                             region == 'UNREGULATED' ~ total_hours + hours_in_arg_gap$total_hours,
                                             TRUE ~ total_hours),
           total_fishing_hours_adj = case_when(region == 'NATIONAL' ~ total_fishing_hours - hours_in_arg_gap$total_fishing_hours,
                                                region == 'UNREGULATED' ~ total_fishing_hours + hours_in_arg_gap$total_fishing_hours,
                                                TRUE ~ total_fishing_hours)) %>% 
  mutate(frac_total_hour = total_hours/sum(total_hours),
         frac_fishing_hours = total_fishing_hours/sum(total_fishing_hours),
         frac_total_hour_adj = total_hours_adj/sum(total_hours_adj),
         frac_fishing_hours_adj = total_fishing_hours_adj/sum(total_fishing_hours_adj))
```



#### Total Fishing Hours (using Marine Regions)

```{r}
(total_fishing_hours_reg <- squid_regulatory_positions_filtered %>%
   mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  group_by(year, flag, region) %>%
  summarise(total_fishing_hours = sum(regional_fishing_hours, na.rm = TRUE)) %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_),
         region = case_when(region == 'NATIONAL' ~ 'EEZ',
                            TRUE ~ region)) %>%
  ggplot() +
  geom_col(aes(year, total_fishing_hours, fill = flag)) +
  #scale_fill_manual('', values = fishwatchr::gfw_palettes$chart[c(TRUE, FALSE)]) +
  scale_fill_manual('',values = c('#EA67B7', '#5FCD8D', '#EFAB68', '#5C98D0')) +
  scale_y_continuous('Total Fishing Hours', limits = c(0, 1250000)) +
  theme_gfw() +
  theme(plot.title.position = 'plot',
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.background = element_rect(fill = '#ffffff'),
        panel.background = element_rect(fill = '#ffffff'),
        axis.title.x = element_blank(),
        plot.caption = element_text(size = 7, hjust = 1)) +
  facet_wrap(~ region))
```


```{r}
save_plots(total_fishing_hours_reg, 
           file_name = glue::glue('total_fishing_hours_reg_{figure_version}'),
           plot_width = 200,
           plot_height = 80,
           units = 'mm',
           dpi = 600,
           output_type = 'png',
           figure_version = figure_version)
```


# Have fishing hours been increasing over time?
## IN SUPPLEMENT
```{r}
(total_fishing_hours_by_reg_region <- squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  group_by(year, region) %>%
  summarise(total_fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop') %>%
  mutate(region = case_when(region == 'NATIONAL' ~ 'EEZ',
                            TRUE ~ region)) %>%
  ggplot() +
  geom_col(aes(year, total_fishing_hours, fill = region)) +
  #scale_fill_manual('', values = fishwatchr::gfw_palettes$chart[c(1,5,7)]) +
  scale_fill_manual('', values = c('#204280','#F8BA47','#EE6256')) + 
  scale_y_continuous('Total Fishing Hours') +
  theme_gfw() +
  theme(plot.title.position = 'plot',
        axis.title.x = element_blank(),
        legend.position = 'bottom',
        plot.background = element_rect(fill = '#ffffff'),
        panel.background = element_rect(fill = '#ffffff'),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.caption = element_text(size = 7, hjust = 1))
)
```

```{r}
save_plots(total_fishing_hours_by_reg_region, 
           file_name = glue::glue('total_fishing_hours_reg_stacked_{figure_version}'),
           plot_width = 120,
           plot_height = 120,
           units = 'mm',
           dpi = 600, 
           output_type = 'png',
           figure_version = figure_version)
```



#### Fraction of Time Spent in Different Regulatory Regimes

```{r}
(fraction_fish_hrs_region <- squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  dplyr::select(year, flag, area, region, regional_fishing_hours) %>%
  mutate(region = case_when(region == 'NATIONAL' & flag == area ~ 'Domestic EEZ',
                            region == 'NATIONAL' & flag != area ~ 'Foreign EEZ',
                            region == 'NATIONAL' & is.na(area) ~ NA_character_,
                            TRUE ~ region)) %>%
  filter(!is.na(region)) %>%
  group_by(year, flag, region) %>%
  summarize(hours = sum(regional_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  inner_join(squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  dplyr::select(year, flag, regional_fishing_hours) %>%
  group_by(year, flag) %>%
  summarize(total_hours = sum(regional_fishing_hours, na.rm = TRUE)) %>%
  ungroup(), 
  by = c('year' = 'year','flag' = 'flag')) %>%
  mutate(frac = hours/total_hours) %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_),
                  region = case_when(region == 'Domestic EEZ' ~ 'DOMESTIC EEZ',
                            region == 'Foreign EEZ' ~ 'FOREIGN EEZ',
                            TRUE ~ region)) %>%
  ggplot() +
  geom_col(aes(year, frac, fill = region)) +
  scale_fill_manual('', values = c('#204280','#7277A4','#F8BA47','#EE6256')) +
  #scale_fill_manual('', values = c("#204280", "#5d69B1", "#ee6256", "#f8ba47")) +
  scale_y_continuous('Fraction of Fishing Hours') +
  theme_gfw() +
  theme(plot.title.position = 'plot',
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.background = element_rect(fill = '#ffffff'),
        panel.background = element_rect(fill = '#ffffff'),
        axis.title.x = element_blank(),
        legend.position = 'bottom',
        plot.caption = element_text(size = 7, hjust = 1)) +
  facet_wrap(~flag)
  )




table_of_data <- squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  dplyr::select(year, flag, area, region, regional_fishing_hours) %>%
  mutate(region = case_when(region == 'NATIONAL' & flag == area ~ 'Domestic EEZ',
                            region == 'NATIONAL' & flag != area ~ 'Foreign EEZ',
                            region == 'NATIONAL' & is.na(area) ~ NA_character_,
                            TRUE ~ region)) %>%
  filter(!is.na(region)) %>%
  group_by(year, flag, region) %>%
  summarize(hours = sum(regional_fishing_hours, na.rm = TRUE)) %>%
  ungroup() %>%
  inner_join(squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021) %>%
  dplyr::select(year, flag, regional_fishing_hours) %>%
  group_by(year, flag) %>%
  summarize(total_hours = sum(regional_fishing_hours, na.rm = TRUE)) %>%
  ungroup(), 
  by = c('year' = 'year','flag' = 'flag')) %>%
  mutate(frac = hours/total_hours) %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_),
        region = case_when(region == 'Domestic EEZ' ~ 'DOMESTIC EEZ',
                            region == 'Foreign EEZ' ~ 'FOREIGN EEZ',
                            TRUE ~ region)) %>%
  filter(region %in% c('DOMESTIC EEZ','FOREIGN EEZ'))

```


```{r}
save_plots(fraction_fish_hrs_region, 
           file_name = glue::glue('frac_fishing_hours_reg_national_eez_{figure_version}'),
           plot_width = 125,
           plot_height = 130,
           units = 'mm',
           dpi = 600, 
           output_type = 'png',
           figure_version = figure_version)
```


# Top 4 EEZs represent 99% of the fishing in EEZs
```{r}
squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021 & 
         flag %in% c('CHN','KOR','TWN','JPN') & 
         region == 'NATIONAL') %>%
  group_by(flag, area,.drop = TRUE) %>%
  summarise(fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop') %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_)) %>%
  group_by(area, .drop = TRUE) %>%
  summarize(total_hours = sum(fishing_hours)) %>%
  mutate(total = sum(total_hours),
         fraction = total_hours/total) %>%
  arrange(desc(fraction)) %>%
  mutate(cumulative_sum = cumsum(fraction))
```




```{r}

missing_levels <- tibble::tibble(flag = c('Japan','Japan'),
                                 area = c('ARG','FLK'),
                                 fishing_hours = c(0,0))


(fishing_in_eezs_byflag <- squid_regulatory_positions_filtered %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  filter(year < 2021 & 
         flag %in% c('CHN','KOR','TWN','JPN') & 
         region == 'NATIONAL') %>%
  group_by(flag, area) %>%
  summarise(fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop') %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_)) %>%
  filter(fishing_hours > 24 & area %in% c('FLK','ARG','JPN','RUS')) %>%
  rbind(missing_levels) %>%
  mutate(area = case_when(area == 'FLK' ~ 'Falkland Islands/Malvinas',
                          area == 'ARG' ~ 'Argentina',
                          area == 'JPN' ~ 'Japan',
                          area == 'RUS' ~ 'Russia',
                          TRUE ~ NA_character_)) %>%
ggplot() +
  geom_col(aes(area, fishing_hours, fill = flag), position = position_dodge()) +
  scale_fill_manual('',values = c('#EA67B7', '#5FCD8D', '#EFAB68', '#5C98D0')) +
  #geom_hline(yintercept = 5263) + 
  scale_x_discrete('') +
  scale_y_continuous('Fishing Hours') +
  theme_gfw() +
  theme(panel.background = element_rect(fill = '#ffffff'),
        plot.background = element_rect(fill = '#ffffff'),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = 'bottom')
)

```


```{r}
save_plots(fishing_in_eezs_byflag, 
           file_name = glue::glue('fishing_hours_ineez_byflag_{figure_version}'),
           output_type = '.png',
           plot_width = 160,
           plot_height = 130,
           units = 'mm',
           dpi = 300,
           figure_version = figure_version)
```




```{r}
squid_regulatory_positions <- fishwatchr::gfw_query(
  'SELECT * FROM `paper_global_squid.squid_regulatory_positions_eez_v20220525`',
  save_query = FALSE,
  run_query = TRUE,
  con = con
)$data

squid_arg_diff_positions <- fishwatchr::gfw_query(
  'SELECT * FROM `paper_global_squid.squid_argentina_diff_positions_v20220525`',
  save_query = FALSE,
  run_query = TRUE,
  con = con
)$data
```

```{r}
squid_regulatory_positions_filtered <- squid_regulatory_positions %>%
  filter(flag %in% c('JPN','KOR','TWN','CHN')) %>%
  filter(region %in% c('NATIONAL','REGULATED','UNREGULATED'))

```






```{r}
missing_levels <- tibble::tibble(flag = c('Japan','Japan'),
                                 area = c('ARG','FLK'),
                                 fishing_hours = c(0,0))

(fishing_in_eezs_byflag_MR_eez <- squid_regulatory_positions_filtered %>%
   filter(!area %in% c('nw_indian','nw_pacific','se_pacific','sw_atlantic')) %>%
  group_by(flag, area) %>%
  summarise(fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop') %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_)) %>%
  filter(fishing_hours > 24 & area %in% c('FLK','ARG','JPN','RUS')) %>%
  rbind(missing_levels) %>%  
  mutate(area = case_when(area == 'FLK' ~ 'Falkland Islands/Malvinas',
                          area == 'ARG' ~ 'Argentina',
                          area == 'JPN' ~ 'Japan',
                          area == 'RUS' ~ 'Russia',
                          TRUE ~ NA_character_)) %>%
ggplot() +
  geom_col(aes(area, fishing_hours, fill = flag), position = position_dodge()) +
  scale_fill_manual('',values = c('#EA67B7', '#5FCD8D', '#EFAB68', '#5C98D0')) +
  #geom_hline(yintercept = 5263) + 
  scale_x_discrete('') +
  scale_y_continuous('Fishing Hours') +
  theme_gfw() +
  theme(panel.background = element_rect(fill = '#ffffff'),
        plot.background = element_rect(fill = '#ffffff'),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank())
)
#China in Argentina 13545.17889
# Republic of Korea	4584.85000
# Taiwan 14969.55694
#
```





# Correcting for Argentina boundary (using Argentina boundary)
```{r}
(fishing_in_eezs_byflag_ARG_eez <- squid_regulatory_positions_filtered %>%
  mutate(year = format(date, '%Y')) %>%
  filter(!area %in% c('nw_indian','nw_pacific','se_pacific','sw_atlantic')) %>%
  filter(year < 2021) %>%
  group_by(flag, area) %>%
  summarise(fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop') %>%
  left_join(squid_arg_diff_positions %>%
  mutate(year = format(date, '%Y')) %>%
  mutate(area = rep('ARG', nrow(.))) %>%
  filter(flag %in% c('CHN','JPN','TWN','KOR')) %>%
  filter(year < 2021) %>%
  group_by(flag, area) %>%
  summarise(total_fishing_hours = sum(regional_fishing_hours, na.rm = TRUE), .groups = 'drop'), by = c('flag')) %>%
  mutate(regional_fishing_hours = case_when(area.x == 'ARG' & area.y == 'ARG' ~ fishing_hours - total_fishing_hours,
                                 #region.x == 'UNREGULATED' & region.y == 'UNREGULATED' ~ total_fishing_hours.x + total_fishing_hours.y,
                                 TRUE ~ fishing_hours)) %>%
  mutate(flag = case_when(flag == 'JPN' ~ 'Japan',
                          flag == 'TWN' ~ 'Chinese Taipei',
                          flag == 'CHN' ~ 'China',
                          flag == 'KOR' ~ 'Republic of Korea',
                          TRUE ~ NA_character_)) %>%
  filter(fishing_hours > 24 & area.x %in% c('FLK','ARG','JPN','RUS')) %>%
  select(flag, area = area.x,  fishing_hours = regional_fishing_hours) %>%
  rbind(missing_levels) %>%  
  mutate(area = case_when(area == 'FLK' ~ 'Falkland Islands/Malvinas',
                          area == 'ARG' ~ 'Argentina',
                          area == 'JPN' ~ 'Japan',
                          area == 'RUS' ~ 'Russia',
                          TRUE ~ NA_character_)) %>%
  mutate(fishing_hours = ifelse(fishing_hours < 0, 0, fishing_hours)) %>%
ggplot() +
  geom_col(aes(area, fishing_hours, fill = flag), position = position_dodge()) +
  scale_fill_manual('',values = c('#EA67B7', '#5FCD8D', '#EFAB68', '#5C98D0')) +
  scale_x_discrete('') +
  scale_y_continuous('Fishing Hours') +
  theme_gfw() +
  theme(panel.background = element_rect(fill = '#ffffff'),
        plot.background = element_rect(fill = '#ffffff'),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank())
)
# China in Argentina -903.625833
# Korea -1.628611
#Taiwan  111.380000
111.38/14969.55694
```


```{r}
(fishing_in_eez_final <- fishing_in_eezs_byflag_MR_eez +  fishing_in_eezs_byflag_ARG_eez + patchwork::plot_layout(ncol = 2, nrow = 1, guides = 'collect') + patchwork::plot_annotation(tag_levels = 'A'))
```


```{r}
save_plots(fishing_in_eez_final, 
           file_name = glue::glue('fishing_in_eezs_byflag_ARG_{figure_version}'),
           output_type = 'png',
           plot_width = 250,
           plot_height = 120,
           units = 'mm',
           dpi = 600,
           figure_version = figure_version)
```

