---
title: "New Vessels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(fishwatchr)
```

## Load data in `00_data_startup`  
  
Load this dataset  
`squid_fleet_connections_2`


```{r}
new_in_year_f <- function(year_list, year) {
  
  if (year == 2018) {

  if (( year %in% year_list) & !((year - 1) %in% year_list)) {
    TRUE
  } else {
    FALSE
  }
  } else if (year == 2019) {
    
  if (( year %in% year_list) & !(2017 %in% year_list) & !(2018 %in% year_list)) {
    TRUE
  } else {
    FALSE
  }
  } else if (year == 2020) {
    
  if (( year %in% year_list) & !(2017 %in% year_list) & !(2018 %in% year_list) & !(2019 %in% year_list)) {
    TRUE
  } else {
    FALSE
  }
  }
}

old_in_year_f <- function(year_list, year) {
  if (year == 2018) {
    if ((year %in% year_list) & (2017 %in% year_list)) {
      TRUE
    } else {
      FALSE
    }
  } else if (year == 2019) {
    if ((year %in% year_list) & ((2017 %in% year_list) | (2018 %in% year_list))) {
      TRUE
    } else {
      FALSE
    }
  } else if (year == 2020) {
    if ((year %in% year_list) & ((2017 %in% year_list) | (2018 %in% year_list) | (2019 %in% year_list))) {
      TRUE
    } else {
      FALSE
    }
  }
}

```

```{r}
number_in_2017 <- squid_fleet_connections_2 %>%
  filter(flag %in% c('CHN','JPN','KOR','TWN')) %>%
  filter(year == 2017) %>%
  group_by(ssvid) %>%
  count(.) %>%
  filter(n > 5) %>%
  distinct(ssvid) %>%
  nrow()

# 794 vessels in 2017
```




```{r}
new_vessels_count <- squid_fleet_connections_2 %>%
  filter(flag %in% c('CHN','JPN','KOR','TWN')) %>%
  #filter(!ssvid %in% inactive_vessels) %>%  #### NOTE REMOVAL OF INACTIVE VESSELS
  group_by(ssvid, year) %>%
  count(.) %>%
  filter(n > 5) %>%
  distinct(.) %>%
  group_by(ssvid) %>%
  summarize(years = list(year)) %>%
  mutate(new_in_2018 = map_lgl(.x = years, .f = new_in_year_f, year = 2018),
         new_in_2019 = map_lgl(.x = years, .f = new_in_year_f, year = 2019),
         new_in_2020 = map_lgl(.x = years, .f = new_in_year_f, year = 2020),
         old_in_2018 = map_lgl(.x = years, .f = old_in_year_f, year = 2018),
         old_in_2019 = map_lgl(.x = years, .f = old_in_year_f, year = 2019),
         old_in_2020 = map_lgl(.x = years, .f = old_in_year_f, year = 2020)) %>%
  ungroup()
```




How many vessels
## Note filtered out CHN, JPN, KOR, TWN
```{r}
new_2018 <- new_vessels_count %>% filter(new_in_2018 == TRUE) %>% nrow() # 165
new_2019 <- new_vessels_count %>% filter(new_in_2019 == TRUE) %>% nrow() # 198
new_2020 <- new_vessels_count %>% filter(new_in_2020 == TRUE) %>% nrow() # 133

existing_2018 <- new_vessels_count %>% filter(old_in_2018 == TRUE) %>% nrow() # 718
existing_2019 <- new_vessels_count %>% filter(old_in_2019 == TRUE) %>% nrow() # 782
existing_2020 <- new_vessels_count %>% filter(old_in_2020 == TRUE) %>% nrow() # 873

number_in_2017 # (from above)
y <- existing_2018  + new_2018 
x <- y - existing_2019
z <- existing_2019 + new_2019 
w <- z - existing_2020
v <- existing_2020 + new_2020


# 794 in 2017
# In 2018: existing = 718
# 718 + 165  = 883
# lost: 883 - 782 = 101
# 782 + 198 = 980
# lost: 980 - 873 = 107
# 873 + 133 = 1006
```



In 2018 there were 
165 new vessels (not seen in 2017), 
198 in 2019 (not seen in 2017 and 2018) and 
133 in 2020 (not seen in 2017, 2018, 2019)

In contrast, there were
718 vessels in 2018 that also existed in 2017
782 vessel in 2019 also seen in (2017 and 2018) and
873 vessels in 2020 also seen in (2017, 2018, 2019)

In 2018 there were 883 vessels (165 were gained)
In 2019 there were 980 vessels (198 were gained and 101 were lost)
In 2020 there were 1006 vessels (133 were gained and 107 were lost)

It appears that there is turnover in the fleet, but a general increase in the number of vessels


```{r}
df <- tibble::tibble(year = c(2017,rep(2018, 3), rep(2019, 3), rep(2020, 3)),
                     count = c(number_in_2017, new_2018, existing_2018, existing_2018-number_in_2017, new_2019, existing_2019, -x, new_2020, existing_2020, -w ),
                     state = c('retained',rep(c('gained', 'retained', 'lost'), 3))) %>%
  mutate(state = factor(state, levels = c('gained','retained','lost')))
```



### All regions at once
```{r}
(new_vessels_bar <- ggplot() +
  geom_col(data = df, aes(year, count, fill = state)) +
  scale_y_continuous('Number of MMSI') + 
  scale_fill_manual('',values = c('#A1D73B','#E6E7EB','#D73BD6')) +
  theme_gfw() +
  theme(plot.title.position = 'plot',
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.background = element_rect(fill = '#ffffff'),
        panel.background = element_rect(fill = '#ffffff'),
        axis.title.x = element_blank()) 
)
```
